---
title: "Simulating animal sound and duty cycles"
author: "A Johnston"
date: "2023.01.25"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE)

# -------------------------------------------------------
# package loading

library(tidyverse)
library(RColorBrewer)

# -------------------------------------------------------
# folder paths

wd <- "/Users/Alison/Documents/REPOS/2022_duty_cycle_simulation/"
setwd(wd)


# -------------------------------------------------------
# load in functions

source(paste0(wd, "code/step1_simulation_functions.R"))

```

## Code

All code is in: 

<https://github.com/ali-johnston/2022_duty_cycle_simulation>


## Goal

To simulate 

 - patterns of animal sound across a day
 
 - duty cycles across a day

In order to identify impacts of probability of detecting an animal with a specific duty cycle


## Simulating animal sounds

We want to be able to vary animal sound by:

 - How much sound they make during a 24 hour period
 
 - How that sound is distributed over the day
 
 - Is there a chunk of the day when they are completely quiet? 
 
 - During the day when they could make sound, is the sound distributed evenly or clustered?


## Simulating duty cycles

We want to be able to vary duty cycles by:

 - How much recording within a 24 hour period
 - How that recording is distributed in chunks of minutes/hours


## Simulations

The next section will show different components of the simulation in sequence. 


## Intro to animal sound simulation {.smaller}

This is one simulation of sounds across a 24 hour period for 5 animals: 

```{r, out.width = "70%", out.height = "20%"}
# add clustering of sounds
par(mfrow = c(1,1), mar = c(0, 5, 0, 1))
test_sound1 <- sim_sound(p_sound = 0.20,		
					p_silence_consec = 0.20, 	
					number_units = 480, 		
					cluster_strength = 0,		# create clustering of sounds within sound window
					random_start = FALSE, 		
					n_animals = 5, 
					order = 61)

```

* The day is broken up into 480 x 3-minute chunks of time

* Within each 3-minute period the animal is either making sound (blue) or is quiet (white)

* Each animal makes sound for 20% of the 3-minute chunks. So 96 of the 480 chunks in their day are blue

* Each animal is completely silent for 96 consecutive chunks at the end of the day. 

* So the 96 sound chunks are distributed randomly among the 384 chunks at the start of the day

* All of these parameters can be altered


## We can alter the amount of time in the day that a species is consecutively quiet

(the number of consecutive white chunks at the end of each row)

```{r, out.width = "80%", out.height = "50%"}
par(mfrow = c(2, 1), mar = c(2, 5, 4, 1))
test_sound1 <- sim_sound(p_sound = 0.20,		# 20% of chunks with sound occuring
					p_silence_consec = 0.70, 	# consecutive 80% of each day with no sound
					number_units = 480, 		# 3 minute chunks. 480 over 24-hour period
					cluster_strength = 0,
					random_start = FALSE, 		# don't randomise start (easier to see what it's doing)
					n_animals = 5, 
					order = 61, 
					main = "70% consecutive quiet")

# add clustering of sounds
test_sound2 <- sim_sound(p_sound = 0.20,		
					p_silence_consec = 0.10, 	
					number_units = 480, 		
					cluster_strength = 0,		# create clustering of sounds within sound window
					random_start = FALSE, 		
					n_animals = 5, 
					order = 61, 
					main = "10% consecutive quiet")

```


## We can alter the amount of time during the day each species is making sound 

(the number of blue chunks in each row)

```{r, out.width = "80%", out.height = "50%"}
par(mfrow = c(2, 1), mar = c(2, 5, 4, 1))
test_sound1 <- sim_sound(p_sound = 0.10,		# 20% of chunks with sound occuring
					p_silence_consec = 0.50, 	# consecutive 50% of each day with no sound
					number_units = 480, 		# 3 minute chunks. 480 over 24-hour period
					cluster_strength = 0,
					random_start = FALSE, 		# don't randomise start (easier to see what it's doing)
					n_animals = 5, 
					order = 61, 
					main = "10% of day with sound")

test_sound2 <- sim_sound(p_sound = 0.40,		
					p_silence_consec = 0.50, 	
					number_units = 480, 		
					cluster_strength = 0,		# create clustering of sounds within sound window
					random_start = FALSE, 		
					n_animals = 5, 
					order = 61, 
					main = "40% of day with sound")

```


## We can alter the clustering of sound during the sound period 

```{r, out.width = "80%", out.height = "50%"}
par(mfrow = c(2, 1), mar = c(2, 5, 4, 1))
test_sound1 <- sim_sound(p_sound = 0.30,		# 20% of chunks with sound occuring
					p_silence_consec = 0.20, 	# consecutive 50% of each day with no sound
					number_units = 480, 		# 3 minute chunks. 480 over 24-hour period
					cluster_strength = NA,
					random_start = FALSE, 		# don't randomise start (easier to see what it's doing)
					n_animals = 5, 
					order = 61, 
					main = "consistent start time")

test_sound2 <- sim_sound(p_sound = 0.30,		
					p_silence_consec = 0.20, 	
					number_units = 480, 		
					cluster_strength = 2,		# create clustering of sounds within sound window
					random_start = FALSE, 		
					n_animals = 5, 
					order = 61, 
					main = "random start time")

```


## We can alter the time during the day when species make sound 


```{r, out.width = "80%", out.height = "50%"}
par(mfrow = c(2, 1), mar = c(2, 5, 4, 1))
test_sound1 <- sim_sound(p_sound = 0.20,		# 20% of chunks with sound occuring
					p_silence_consec = 0.20, 	# consecutive 50% of each day with no sound
					number_units = 480, 		# 3 minute chunks. 480 over 24-hour period
					cluster_strength = NA,
					random_start = FALSE, 		# don't randomise start (easier to see what it's doing)
					n_animals = 5, 
					order = 61, 
					main = "no clustering")

test_sound2 <- sim_sound(p_sound = 0.20,		
					p_silence_consec = 0.20, 	
					number_units = 480, 		
					cluster_strength = NA,		# create clustering of sounds within sound window
					random_start = TRUE, 		
					n_animals = 5, 
					order = 61, 
					main = "strong clustering")

```


## We can alter the number of animals

```{r, out.width = "80%", out.height = "70%"}
par(mfrow = c(1, 1), mar = c(2, 5, 4, 1))
test_sound1 <- sim_sound(p_sound = 0.20,		# 20% of chunks with sound occuring
					p_silence_consec = 0.50, 	# consecutive 50% of each day with no sound
					number_units = 480, 		# 3 minute chunks. 480 over 24-hour period
					cluster_strength = 2,
					random_start = TRUE, 		# don't randomise start (easier to see what it's doing)
					n_animals = 100, 
					order = 61, 
					main = "100 animals")

```


## Adding duty cycles


## There are also functions to simulate duty cycles

 - These alter the amount of time during a day that recording is happening
 - And how recording is distributed over the day
 
 
## Simulating different duty cycles with animals

(red = recording on)

```{r, out.width = "80%", out.height = "50%"}
par(mfrow = c(2, 1), mar = c(2, 5, 4, 1))

test_sound_duty1 <- sim_sound_duty(p_sound = 0.10,
					p_silence_consec = 0.70,
					number_units = 480,
					cluster_strength = NA,
					order = 61,
					random_start_sound = TRUE, 
					n_animals = 5, 
					prop_recording = 0.20,
					units_per_chunk = 40,
					random_start_duty = TRUE,
					plot_sound_duty = TRUE)

test_sound_duty2 <- sim_sound_duty(p_sound = 0.10,
					p_silence_consec = 0.70,
					number_units = 480,
					cluster_strength = 2,
					order = 61,
					random_start_sound = TRUE, 
					n_animals = 5, 
					prop_recording = 0.20,
					units_per_chunk = 40,
					random_start_duty = TRUE,
					plot_sound_duty = TRUE)

```


## Summarising animals 'caught'

```{r, eval = TRUE, echo = TRUE}
sim1 <- sim_sound_duty(
  p_sound = 0.10,          # proportion of sound 
  p_silence_consec = 0.70, # prop of consecutive silence
  number_units = 480,      # number of time units 
  cluster_strength = NA,   # cluster strength (NA = none)
  random_start_sound = TRUE, # random start time of sound
  n_animals = 1000,        # number of animals
  prop_recording = 0.20,   # prop of day with recording
  units_per_chunk = 40,    # no. time units per recording
  random_start_duty = TRUE,# random start time of DC
  plot_sound_duty = FALSE, # plot the simulations
  seed = 283)              # set.seed for simulation
```


## Summarising animals 'caught'

```{r, echo = TRUE, eval = FALSE}

hist(sim1$no_detections)

```


```{r, eval = TRUE, echo = FALSE}

hist(sim1$no_detections, breaks = seq(0, 31, by = 1), col = c(alpha("firebrick", 0.4), rep("grey", 30)), main = "")

# proportion animals not detected
round(mean(sim1$no_detections == 0), 2)
```


## Simulating an array of parameters 

```{r, echo = FALSE, eval = TRUE, cache = TRUE, out.height = "80%", out.width = "75%"}
# for now the simulations are in units of 3 mins

# set up variation in proportion of the day with sound
single_unit_mins <- 3
tot_mins <- 24*60
tot_units <- tot_mins / single_unit_mins
units_sound <- seq(1, round(tot_units*0.25), by = 5)

# set up variation in how many units of consecutive silence
units_silence_consec <- seq(0.5, 0.95, by = 0.05)*tot_units |> round()
units_sound_contained <- tot_units - units_silence_consec

# set up distribution in gaps between units with sound



#########################################################
## set duty cycle parameters 

## for now just picking one as a test case
## eventually this will be automated in an array of 
## simulations

propn_sampling <- 0.25
number_units_per_chunk <- 40 
# with 3-minute units:
# 20 units = 1 hour



#########################################################
## set up results matrix and loop through simulations

res_mat <- matrix(NA, nrow = length(units_sound), ncol = length(units_silence_consec))

for(kkk in 1:length(units_sound)){

	viable_gaps <- (tot_units - units_silence_consec) >= units_sound[kkk]
	if(any(viable_gaps)){
		for(jjj in which(viable_gaps)){

			n_animals <- 100

			run_sim <- sim_sound_duty(p_sound = units_sound[kkk] / tot_units,
								p_silence_consec = units_silence_consec[jjj] / tot_units,
								number_units = tot_units,
								cluster_strength = 1,
								seed = (kkk*1000) + jjj, 
								n_animals = n_animals, 
								order = 41,
								prop_recording = propn_sampling,
								units_per_chunk = number_units_per_chunk)

			detected <- run_sim$no_detections > 0
			res_mat[kkk, jjj] <- as.numeric(sum(detected) / length(detected))

		}
	}
}


res_long <- data.frame(units_sound = rep(units_sound, length(units_silence_consec)),
						units_silence_consec = sort(rep(units_silence_consec, length(units_sound))),
						prop_detected = as.vector(res_mat)) |>
				mutate(prop_sound = units_sound / tot_units) |>
				mutate(prop_sound_contained = 1 - (units_silence_consec / tot_units))


ggplot(data=res_long, aes(x = prop_sound, y = prop_sound_contained)) + 	
	geom_tile(aes(fill = prop_detected)) + 
	scale_fill_gradientn(limits = c(0, 1), colours = brewer.pal(8, "YlGnBu")) + 
	xlab("Proportion of day with sound") + 
	ylab("Proportion of day containing all sound") + 
	theme_bw() + 
	theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())


```





